How to select predictive models for decision-making or causal inference?

### 2024-06-14 how_to_select.tex
manuscript for gigascience: Rename images to match jounral <=60 chars

Prediction models have long been used in medicine, as with risk
score or prognostic models \cite{moons2009prognosis,steyerberg2019clinical}.
While these have historically been simple models on simple data, this is
changing with progress in machine learning and richer medical data
\cite{beam2018big,rajkomar2019machine}. Health predictions can now integrate medical images
\cite{khojaste2022deep,zhang2019radiological,yala2019deep,shen2019deep,nassif2022breast},
patient records
\cite{mooney2018bigdata,desai2020comparison,simon2018predicting} or clinical
notes \cite{horng2017creating,wang2020prediction,spasic2020clinical}.
Complex data is difficult to control and model, but these models are
validated by verifying the accuracy of the prediction on left-out data
\cite{altman2009prognosis,poldrack2020establishment,varoquaux2022evaluating}.
Crucial to the clinical adoption of a model predicting a health outcome
is that it ``can support decisions about patient care''
\cite{wyatt1995commentary}. Precision medicine is about
guiding decisions: \emph{eg} will an individual benefit from an intervention such as surgery
\cite{fontana2019can}? An estimate of the effect of the treatment can be
obtained by contrasting model predictions with and without the treatment,
but statistical validity requires causal inference
\cite{snowden_implementation_2011,sperrin2019explicit,blakely2020reflection}.

@@

Indeed, concluding on the effect of a treatment is a difficult
causal-inference task, as it can be easily compromised by confounding:
spurious associations between treatment
allocation and baseline health, \eg~only prescribing a drug to mild cases
\cite{hernan_causal_2020,vanderweele2019principles}.
Predictive modeling bridges to causal inference theory under the name of
\emph{outcome models} (or G-computation,
G-formula \cite{robins_role_1986}, Q-model
\cite{snowden_implementation_2011}, conditional mean regression
\cite{wendling_comparing_2018}).
% Controlled allocation of treatment, as in randomized controlled trials (RCTs),
% alleviate this concern. Yet most machine-learning models are trained on
% \emph{observational} data, close to real-word practice
% \cite{black1996we,hernan_methods_2021} but challenging for causal inference.
Medical statistics and epidemiology have mostly used other
causal-inference methods, modeling treatment assignment
with propensity scores \cite{rosenbaum_central_1983,austin_moving_2015,casucci2018estimating,grose_use_2020}. Outcome modeling brings
the benefit of going
beyond average effects, estimating individualized or conditional average
treatment effects (CATE), central to precision medicine.
%
For this purpose, such methods are also invaluable on randomized trials
\cite{su2018random,lamont2018identification,hoogland2021tutorial}.

@@

\begin{figure}[!b]
    % XXX: for submission
    \centering\begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{_1_r_risk_domination_r_risk_domination__ref_metric_mean_risks_by_Dataset_feasible_only.pdf}
    \end{minipage}
    \caption{\textbf{The $R$-risk is the best metric}: Relative Kendall's $\tau$ agreement with $\tau\text{-risk}$.
        Strong and Weak overlap correspond to the first and last tertiles of the overlap distribution measured with
        Normalized Total Variation eq. \ref{eq:ntv}. \ref{apd:experiments:additional_results} presents the same results
        by adding semi-oracle risks in Figure \ref{apd:fig:relative_kendalls_all_datasets_all_metrics}, measured with
        absolute Kendall's in Figure \ref{apd:fig:all_datasets_tau_risk_ranking_agreement} and with $\tau\mathrm{-risk}$
        gains in Figure \ref{apd:all_datasets_normalized_bias_tau_risk_to_best_method}. Table
        \ref{apd:table:relative_kendalls_all_datasets} gives median and
        IQR of the relative Kendall.}\label{fig:relative_kendalls_all_datasets}
\end{figure}

@@

\begin{figure}[!h]
    % XXX: for submission
    \centering\begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{_2_overlap_influence_overlap_by_bin_comparaison_kendall_by_Dataset_r_risk_only_twocols.pdf}
    \end{minipage}
    \caption{\textbf{Model selection is harder for low population
            overlap}:
        Kendall's $\tau$ agreement with $\tau\text{-risk}$. Strong, medium and Weak overlap
        are the tertiles of the overlap measured with NTV eq. \ref{eq:ntv}. Supplementary materials presents results for all
        metrics in Figure \ref{apd:fig:all_datasets_overlap_effect} in absolute
        Kendall's and continuous overlap values in Figure
            {\ref{apd:fig:all_datasets_tau_risk_ranking_agreement}}.}\label{fig:all_datasets_overlap_effect_r_risk}
\end{figure}

@@

\begin{figure}[!tb]
    % XXX: for submission
    \centering\begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{_4_nuisance_models_r_risk_only_3datasets_twocols.pdf}
    \end{minipage}

@@

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.95\linewidth]{2023-03-08-11-10-28_acic_2016_ate_heterogeneity.parquet_abs_bias_ylog_scale=True.pdf}%
    \caption{\textbf{Different outcome models lead to different
            estimation errors on the Average Treatment Effects},
        on 77 classic simulations with known true causal effect
        \cite{dorie_automated_2019}. The different models are ridge regression
        and random forests with different hyper-parameters
        (details
        \ref{apd:toy_example:acic_2016_ate_variability}). The different configurations are
        plotted as a function of increasing difference between treated and
        untreated population --see
        \autoref{subsec:measuring_overlap}.
        There is no systematic best performer; data-driven model
        selection is important.
        \label{fig:acic_2016_ate_heterogeneity}%
    }
\end{figure}

@@

\begin{figure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \caption{\textbf{Uncalibrated classifiers}}
        \includegraphics[width=\linewidth]{overlap_measure_diff_oracle_ntv_to_n_tv_calibration=False_vs_oracle_n_tv.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \caption{\textbf{Calibrated
                classifiers}}
        \includegraphics[width=\linewidth]{overlap_measure_diff_oracle_ntv_to_n_tv_calibration=True_vs_oracle_n_tv.pdf}
    \end{subfigure}
    \caption{a) Without calibration, estimation of NTV is not trivial even for
        boosting models. b) Calibrated classifiers are able to recover the true
        Normalized Total Variation for all datasets where it is
        available.}\label{apd:overlap:ntv_approximation}
\end{figure}

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=\linewidth]{_1_r_risk_domination_r_risk_domination__ref_metric_mean_risks_by_Dataset.pdf}
    %\hfill
    \caption{\textbf{The $R$-risk is the best metric}: Relative Kendall's $\tau$
        agreement with $\tau\text{-risk}$. Strong and Weak overlap correspond to
        the first and last tertiles of the overlap distribution measured with
        Normalized Total Variation eq. \ref{eq:ntv}.
    }\label{apd:fig:relative_kendalls_all_datasets_all_metrics}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=\textwidth]{kendalls_tau_caussim__nuisance_non_linear__candidates_ridge__overlap_01-247.pdf}
        \label{fig:ranking_agreement_w_tau_risk_caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=\textwidth]{kendalls_tau_acic_2016__nuisance_non_linear__candidates_hist_gradient_boosting__dgp_1-77__rs_1-10.pdf}
        \label{fig:ranking_agreement_tau_risk_acic_2016}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        \centering
        \includegraphics[width=1.5\textwidth]{legend_metrics.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2018}}
        \includegraphics[width=\textwidth]{kendalls_tau_acic_2018__nuisance_stacking__candidates_hist_gradient_boosting__tset_50.pdf}
        \label{fig:ranking_agreement_w_tau_risk_acic_2018}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{TWINS}}
        \includegraphics[width=\textwidth]{kendalls_tau_twins__nuisance_stacking__candidates_hist_gradient_boosting__tset_50__overlap_11-296__rs_0-9_noise10.pdf}
        \label{fig:ranking_agreement_tau_risk_twins}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        ~
    \end{subfigure}
    % \begin{subfigure}[b]{\textwidth}
    %   \centering
    %   \includegraphics[width=0.5\textwidth]{legend_metrics_horizontal.pdf}
    % \end{subfigure}
    \caption{Agreement with $\tau\text{-risk}$ ranking of methods function
        of overlap violation. The lines represent medians, estimated with a
        lowess. The transparent
        bands denote the 5\% and 95\% confidence intervals.}\label{apd:fig:all_datasets_tau_risk_ranking_agreement}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=\textwidth]{normalized_bias_tau_risk_to_best_method_caussim__nuisance_non_linear__candidates_ridge__overlap_01-247.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=\textwidth]{normalized_bias_tau_risk_to_best_method_acic_2016__nuisance_non_linear__candidates_hist_gradient_boosting__dgp_1-77__rs_1-10.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_acic_2016}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        \centering
        \includegraphics[width=1.5\textwidth]{legend_metrics.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2018}}
        \includegraphics[width=\textwidth]{normalized_bias_tau_risk_to_best_method_acic_2018__nuisance_stacking__candidates_hist_gradient_boosting__tset_50.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_acic_2018}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{TWINS}}
        \includegraphics[width=\textwidth]{normalized_bias_tau_risk_to_best_method_twins__nuisance_stacking__candidates_hgb.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_twins}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        ~
    \end{subfigure}
    \caption{Metric performances by normalized tau-risk distance to the best
        method selected with $\tau\text{-risk}$. All nuisances are learned with the
        same estimator stacking gradient boosting and ridge regression. Doted and
        plain lines corresponds to 60\% lowess quantile estimates. This choice of
        quantile allows to see better the oracle metrics lines for which outliers with a value
        of 0 distord the curves.}
    \label
    {apd:all_datasets_normalized_bias_tau_risk_to_best_method}
\end{figure}

@@

\begin{figure}
    \begin{subfigure}[b]{0.9\textwidth}
        %\centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=1.15\textwidth]{_3_procedure_Caussim_N5000__ref_metric_oracle_r_risk_training_procedure.pdf}
        \label{fig:experiments:procedures_comparison:caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=1\textwidth]{_3_procedure_ACIC2016_N4802__ref_metric_oracle_r_risk_training_procedure.pdf}
        \label{fig:experiments:procedures_comparison:acic_2016}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \caption{\textbf{Twins}}
        \includegraphics[width=1\textwidth]{_3_procedure_Twins_N11984__ref_metric_oracle_r_risk_training_procedure.pdf}
        \label{fig:experiments:procedures_comparison:twins}
    \end{subfigure}
    \hfill
    \caption{Results are similar between the \textcolor{MidnightBlue}{Shared
            nuisances/candidate set} and
        the \textcolor{RedOrange}{Separated nuisances set} procedure. The
        experience has not been run on the full metrics for Caussim due to computation costs.}\label
    {apd:fig:procedures_comparison_all_metrics}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{_2_overlap_influence_overlap_by_bin_comparaison_kendall_by_Dataset.pdf}
    \hfill
    \caption{\textbf{Low population overlap hinders causal model selection for all
            metrics}:
        Kendall's $\tau$ agreement with $\tau\text{-risk}$. Strong, medium and Weak overlap
        correspond
        to the tertiles of the overlap distribution measured with Normalized Total
        Varation eq. \ref{eq:ntv}.}\label{apd:fig:all_datasets_overlap_effect}
\end{figure}

@@

\begin{figure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{\textbf{Twins}}
        \includegraphics[width=1\textwidth]{_4_nuisance_models_Twins_N11984__ref_metric_oracle_r_risk_nuisance_models.pdf}
        \label{fig:experiments:nuisance_comparison:twins}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{\textbf{Twins downsampled}}
        \includegraphics[width=1\textwidth]{_4_nuisance_models_Twinsdownsampled_N4794__ref_metric_oracle_r_risk_nuisance_models.pdf}
        \label{fig:experiments:nuisance_comparison:twins_ds}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        %\centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=1.15\textwidth]{_4_nuisance_models_Caussim_N5000__ref_metric_oracle_r_risk_nuisance_models.pdf}
        \label{fig:experiments:nuisance_comparison:caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=1\textwidth]{_4_nuisance_models_ACIC2016_N4802__ref_metric_oracle_r_risk_nuisance_models.pdf}
        \label{fig:experiments:nuisance_comparison:acic_2016}
    \end{subfigure}
    \hfill
    \caption{Learning the nuisances with \textcolor{DarkOrchid}{stacked models} (linear and
        gradient boosting) is important for successful model selection with R-risk.
        For Twins dataset, there is no improvement for \textcolor{DarkOrchid}{stacked models} compared to
        \textcolor{ForestGreen}{linear models} because of the linearity of the propensity model.}\label
    {apd:fig:nuisances_comparison}
\end{figure}

@@

\begin{figure}[!t]
    \begin{minipage}{.5\textwidth}
        \centerline{\textbf{a) CATE estimation error}}
        \includegraphics[width=\linewidth]{_5_train_size_evaluated_metric_r_risk__evaluation_validation_tau_risk__best_tau_norm__False_norm_False__acic16_twocols.pdf}
    \end{minipage}
    %\hspace{0.3\textwidth}
    \begin{minipage}{.5\textwidth}
        \centerline{\textbf{b) ATE estimation error}}
        \includegraphics[width=\linewidth]{_5_train_size_evaluated_metric_r_risk__evaluation_validation_test_abs_bias_ate__best_tau_norm__False_norm_True__acic16_twocols.pdf}
    \end{minipage}%
    \caption{\textbf{a) For CATE, a train/test ratio of 0.9/0.1 appears a good
            trade-off.} b) For ATE, there is a small signal pointing also to
        0.9/0.1 (K=10).
        for ATE. Experiences on 10 replications of all 78 instances of the ACIC 2016
        data.}\label{fig:train_test_ratio}
\end{figure}

@@

M.D. conceived and conducted the experiments, M.D. and G.V. analysed the results. M.D. and G.V. wrote and reviewed the manuscript.

>>>>

Prediction models have long been used in biomedical settings, as with risk
score or prognostic models \cite{moons2009prognosis,steyerberg2019clinical}.
While these have historically been simple models on simple data, this is
changing with progress in machine learning and richer medical data
\cite{beam2018big,rajkomar2019machine}. Health predictions can now integrate medical images
\cite{khojaste2022deep,zhang2019radiological,yala2019deep,shen2019deep,nassif2022breast},
patient records
\cite{mooney2018bigdata,desai2020comparison,simon2018predicting} or clinical
notes \cite{horng2017creating,wang2020prediction,spasic2020clinical}.
Complex data is difficult to control and model, but these models are
validated by verifying the accuracy of the prediction on left-out data
\cite{altman2009prognosis,poldrack2020establishment,varoquaux2022evaluating}.
Crucial to the clinical adoption of a model predicting a health outcome
is that it ``can support decisions about patient care''
\cite{wyatt1995commentary}. Precision medicine is about
guiding decisions: \emph{eg} will an individual benefit from an intervention such as surgery
\cite{fontana2019can}? An estimate of the effect of the treatment can be
obtained by contrasting model predictions with and without the treatment,
but statistical validity requires causal inference
\cite{snowden_implementation_2011,sperrin2019explicit,blakely2020reflection}.

@@

Indeed, concluding on the effect of a treatment is a difficult
causal-inference task, as it can be easily compromised by confounding:
spurious associations between treatment
allocation and baseline health, \eg~only prescribing a drug to mild cases
\cite{hernan_causal_2020,vanderweele2019principles}.
Predictive modeling bridges to causal inference theory under the name of
\emph{outcome models} (or G-computation,
G-formula \cite{robins_role_1986}, Q-model
\cite{snowden_implementation_2011}, conditional mean regression
\cite{wendling_comparing_2018}).
% Controlled allocation of treatment, as in randomized controlled trials (RCTs),
% alleviate this concern. Yet most machine-learning models are trained on
% \emph{observational} data, close to real-word practice
% \cite{black1996we,hernan_methods_2021} but challenging for causal inference.
Medical statistics and epidemiology have mostly used other
causal-inference methods, modeling treatment assignment
with propensity scores \cite{rosenbaum_central_1983,austin_moving_2015,casucci2018estimating,grose_use_2020}. Outcome modeling brings
the benefit of going
beyond average effects, estimating individualized or conditional average
treatment effects (CATE), central to precision medicine.
%
For this purpose, such methods are also invaluable for randomized trials
\cite{su2018random,lamont2018identification,hoogland2021tutorial}.

@@

\begin{figure}[!b]
    % XXX: for submission
    \centering\begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{_1_r_risk_domination_mrisks_by_ds_feasible.pdf}
    \end{minipage}
    \caption{\textbf{The $R$-risk is the best metric}: Relative Kendall's $\tau$ agreement with $\tau\text{-risk}$.
        Strong and Weak overlap correspond to the first and last tertiles of the overlap distribution measured with
        Normalized Total Variation eq. \ref{eq:ntv}. \ref{apd:experiments:additional_results} presents the same results
        by adding semi-oracle risks in Figure \ref{apd:fig:relative_kendalls_all_datasets_all_metrics}, measured with
        absolute Kendall's in Figure \ref{apd:fig:all_datasets_tau_risk_ranking_agreement} and with $\tau\mathrm{-risk}$
        gains in Figure \ref{apd:all_datasets_normalized_bias_tau_risk_to_best_method}. Table
        \ref{apd:table:relative_kendalls_all_datasets} gives median and
        IQR of the relative Kendall.}\label{fig:relative_kendalls_all_datasets}
\end{figure}

@@

\begin{figure}[!h]
    % XXX: for submission
    \centering\begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{_2_overlap_influence_by_bin_k_by_ds_r_risk_2cols.pdf}
    \end{minipage}
    \caption{\textbf{Model selection is harder for low population
            overlap}:
        Kendall's $\tau$ agreement with $\tau\text{-risk}$. Strong, medium and Weak overlap
        are the tertiles of the overlap measured with NTV eq. \ref{eq:ntv}. Supplementary materials presents results for all
        metrics in Figure \ref{apd:fig:all_datasets_overlap_effect} in absolute
        Kendall's and continuous overlap values in Figure
            {\ref{apd:fig:all_datasets_tau_risk_ranking_agreement}}.}\label{fig:all_datasets_overlap_effect_r_risk}
\end{figure}

@@

\begin{figure}[!tb]
    % XXX: for submission
    \centering\begin{minipage}{\linewidth}
        \includegraphics[width=\linewidth]{_4_nuisance_models_r_risk_3ds_2cols.pdf}
    \end{minipage}

@@

M.D. conceived and conducted the experiments, M.D. and G.V. analyzed the results. M.D. and G.V. wrote and reviewed the manuscript.

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=0.95\linewidth]{acic_2016_ate_hte.pdf}%
    \caption{\textbf{Different outcome models lead to different
            estimation errors on the Average Treatment Effects},
        on 77 classic simulations with known true causal effect
        \cite{dorie_automated_2019}. The different models are ridge regression
        and random forests with different hyper-parameters
        (details
        \ref{apd:toy_example:acic_2016_ate_variability}). The different configurations are
        plotted as a function of increasing difference between treated and
        untreated population --see
        \autoref{subsec:measuring_overlap}.
        There is no systematic best performer; data-driven model
        selection is important.
        \label{fig:acic_2016_ate_heterogeneity}%
    }
\end{figure}

@@

\begin{figure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \caption{\textbf{Uncalibrated classifiers}}
        \includegraphics[width=\linewidth]{overlap_measure_diff_oracle_ntv_to_pred_ntv.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \caption{\textbf{Calibrated
                classifiers}}
        \includegraphics[width=\linewidth]{overlap_measure_diff_oracle_ntv_to_ntv_wcalibration.pdf}
    \end{subfigure}
    \caption{a) Without calibration, estimation of NTV is not trivial even for
        boosting models. b) Calibrated classifiers are able to recover the true
        Normalized Total Variation for all datasets where it is
        available.}\label{apd:overlap:ntv_approximation}
\end{figure}

@@

\begin{figure}[!b]
    \centering
    \includegraphics[width=\linewidth]{_1_r_risk_domination_mrisks_by_ds.pdf}
    %\hfill
    \caption{\textbf{The $R$-risk is the best metric}: Relative Kendall's $\tau$
        agreement with $\tau\text{-risk}$. Strong and Weak overlap correspond to
        the first and last tertiles of the overlap distribution measured with
        Normalized Total Variation eq. \ref{eq:ntv}.
    }\label{apd:fig:relative_kendalls_all_datasets_all_metrics}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=\textwidth]{ktau_caussim_nuisance_nlin_models_ridge_ov1-247.pdf}
        \label{fig:ranking_agreement_w_tau_risk_caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=\textwidth]{ktau_acic2016_nuisance_nlin_models_hgb_dgp1-77_rs1-10.pdf}
        \label{fig:ranking_agreement_tau_risk_acic_2016}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        \centering
        \includegraphics[width=1.5\textwidth]{legend_metrics.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2018}}
        \includegraphics[width=\textwidth]{ktau_acic2018_nuisance_stacking_models_hgb_tset_50.pdf}
        \label{fig:ranking_agreement_w_tau_risk_acic_2018}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{TWINS}}
        \includegraphics[width=\textwidth]{ktau_twins_nuisance_stacking__models_hgb.pdf}
        \label{fig:ranking_agreement_tau_risk_twins}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        ~
    \end{subfigure}
    % \begin{subfigure}[b]{\textwidth}
    %   \centering
    %   \includegraphics[width=0.5\textwidth]{legend_metrics_horizontal.pdf}
    % \end{subfigure}
    \caption{Agreement with $\tau\text{-risk}$ ranking of methods function
        of overlap violation. The lines represent medians, estimated with a
        lowess. The transparent
        bands denote the 5\% and 95\% confidence intervals.}\label{apd:fig:all_datasets_tau_risk_ranking_agreement}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=\textwidth]{norm_bias_trisk_tobest_caussim_nuis_nlin_models_ridge.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=\textwidth]{norm_bias_trisk_tobest_acic2016_nuis_nlin_models_hgb.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_acic_2016}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        \centering
        \includegraphics[width=1.5\textwidth]{legend_metrics.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{ACIC 2018}}
        \includegraphics[width=\textwidth]{norm_bias_trisk_tobest_acic2018_nuis_nlin_models_hgb.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_acic_2018}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.44\textwidth}
        \centering
        \caption{\textbf{TWINS}}
        \includegraphics[width=\textwidth]{norm_bias_trisk_tobest_twins_nuis_stack_models_hgb.pdf}
        \label{fig:normalized_bias_tau_risk_to_best_method_twins}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.10\textwidth}
        ~
    \end{subfigure}
    \caption{Metric performances by normalized tau-risk distance to the best
        method selected with $\tau\text{-risk}$. All nuisances are learned with the
        same estimator stacking gradient boosting and ridge regression. Doted and
        plain lines corresponds to 60\% lowess quantile estimates. This choice of
        quantile allows to see better the oracle metrics lines for which outliers with a value
        of 0 distord the curves.}
    \label
    {apd:all_datasets_normalized_bias_tau_risk_to_best_method}
\end{figure}

@@

\begin{figure}
    \begin{subfigure}[b]{0.9\textwidth}
        %\centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=1.15\textwidth]{_3_procedure_Caussim_N5000__ref_metric_oracle_r_risk.pdf}
        \label{fig:experiments:procedures_comparison:caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=1\textwidth]{_3_procedure_ACIC2016_N4802__ref_oracle_r_risk.pdf}
        \label{fig:experiments:procedures_comparison:acic_2016}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \caption{\textbf{Twins}}
        \includegraphics[width=1\textwidth]{_3_procedure_Twins_N11984__ref_oracle_r_risk.pdf}
        \label{fig:experiments:procedures_comparison:twins}
    \end{subfigure}
    \hfill
    \caption{Results are similar between the \textcolor{MidnightBlue}{Shared
            nuisances/candidate set} and
        the \textcolor{RedOrange}{Separated nuisances set} procedure. The
        experience has not been run on the full metrics for Caussim due to computation costs.}\label
    {apd:fig:procedures_comparison_all_metrics}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{_2_overlap_influence_by_bin_k_by_ds.pdf}
    \hfill
    \caption{\textbf{Low population overlap hinders causal model selection for all
            metrics}:
        Kendall's $\tau$ agreement with $\tau\text{-risk}$. Strong, medium and Weak overlap
        correspond
        to the tertiles of the overlap distribution measured with Normalized Total
        Varation eq. \ref{eq:ntv}.}\label{apd:fig:all_datasets_overlap_effect}
\end{figure}

@@

\begin{figure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{\textbf{Twins}}
        \includegraphics[width=1\textwidth]{_4_nuisance_models_Twins_N11984__ref_oracle_r_risk.pdf}
        \label{fig:experiments:nuisance_comparison:twins}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{\textbf{Twins downsampled}}
        \includegraphics[width=1\textwidth]{_4_nuisance_models_Twinssmall_N4794__ref_oracle_r_risk.pdf}
        \label{fig:experiments:nuisance_comparison:twins_ds}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        %\centering
        \caption{\textbf{Caussim}}
        \includegraphics[width=1.15\textwidth]{_4_nuisance_models_Caussim_N5000__ref_oracle_r_risk.pdf}
        \label{fig:experiments:nuisance_comparison:caussim}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{\textbf{ACIC 2016}}
        \includegraphics[width=1\textwidth]{_4_nuisance_models_ACIC2016_N4802__ref_oracle_r_risk.pdf}
        \label{fig:experiments:nuisance_comparison:acic_2016}
    \end{subfigure}
    \hfill
    \caption{Learning the nuisances with \textcolor{DarkOrchid}{stacked models} (linear and
        gradient boosting) is important for successful model selection with R-risk.
        For Twins dataset, there is no improvement for \textcolor{DarkOrchid}{stacked models} compared to
        \textcolor{ForestGreen}{linear models} because of the linearity of the propensity model.}\label
    {apd:fig:nuisances_comparison}
\end{figure}

@@

\begin{figure}[!t]
    \begin{minipage}{.5\textwidth}
        \centerline{\textbf{a) CATE estimation error}}
        \includegraphics[width=\linewidth]{_5_train_size_r_risk__cate__acic16_2cols.pdf}
    \end{minipage}
    %\hspace{0.3\textwidth}
    \begin{minipage}{.5\textwidth}
        \centerline{\textbf{b) ATE estimation error}}
        \includegraphics[width=\linewidth]{_5_train_size_r_risk__ate__acic16_2cols.pdf}
    \end{minipage}%
    \caption{\textbf{a) For CATE, a train/test ratio of 0.9/0.1 appears a good
            trade-off.} b) For ATE, there is a small signal pointing also to
        0.9/0.1 (K=10).
        for ATE. Experiences on 10 replications of all 78 instances of the ACIC 2016
        data.}\label{fig:train_test_ratio}
\end{figure}

@@
